---
title: "ggplot2-template-practice"
author: "Ming Chen"
date: "4/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```


# Geom functions

* geom_point
* geom_line
* geom_point + geom_line
* geom_line + point (ggplot2 is built up by layers, so plotting order matters.)

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_line(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity", color="red") +
  geom_line(mapping = aes(x = disp, y = hp), 
             stat = "identity", size=4) +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_line(mapping = aes(x = disp, y = hp), 
             stat = "identity", size=4) +
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity", color="red") +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```


# Mappings

* x -> x axis
* y -> y axis
* gear -> color (continuous color)
* gear -> color (discrete color)
* am  -> shape (discrete)
* wt -> size (continuous)

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=gear), 
             stat = "identity") +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=as.factor(gear)), 
             stat = "identity") +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=as.factor(gear), shape=as.factor(am)), 
             stat = "identity") +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete() +
  scale_shape_discrete() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=as.factor(gear), 
                           shape=as.factor(am), size=wt),  
             stat = "identity") +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete() +
  scale_shape_discrete() +
  theme()
```


# STAT

```{r}
cat("===== geom_point ======\n")
args(geom_point)
cat("===== geom_line ======\n")
args(geom_line)
cat("===== geom_bar ======\n")
args(geom_bar)
cat("===== geom_histogram ======\n")
args(geom_histogram)
cat("===== geom_bin2d ======\n")
args(geom_bin2d)
cat("===== geom_density ======\n")
args(geom_density)
```

# COORDINATE

* coord_cartesian
* coord_flip
* coord_trans

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_cartesian() +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```


```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_flip() +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_trans(x="log10", y="log10") +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```


# Facet

* facet_null
* split by a variable on y axis
* split by a variable on x axis
* split by multiple variables

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_cartesian() +
  facet_null() +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_cartesian() +
  facet_grid(. ~ am ) +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```


```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_cartesian() +
  facet_grid(am ~ . ) +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```


```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_cartesian() +
  facet_grid(am ~ gear ) +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp), 
             stat = "identity") +
  coord_cartesian() +
  facet_grid(am + gear ~ .) +
  scale_x_continuous() +
  scale_y_continuous() +
  theme()
```

# Scale

* discrete data values -> discreate visual values (x, y, color, shape)
* continuous data values -> continous visual values (x, y, color, size, area, transparency)

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=gear), 
             stat = "identity") +
  coord_cartesian() +
  facet_null() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=as.factor(gear)), 
             stat = "identity") +
  coord_cartesian() +
  facet_null() +
  scale_x_continuous(name = "Displacemen (cu.in.)", 
                     breaks = c(100, 150, 200, 250, 300, 350, 400),
                     labels = c("0.10k", "0.15k", "0.20k", "0.25k", "0.30k", "0.35k", "0.40k")) +
  scale_y_continuous(name = "Gross horsepower", limits = c(0, 500)) +
  scale_color_manual(values = c("3"="red", "4"="blue", "5"="green"),
                     name="Gear") +
  theme()
```

# Theme

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=gear), 
             stat = "identity") +
  coord_cartesian() +
  facet_null() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_continuous() +
  theme()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=gear), 
             stat = "identity") +
  coord_cartesian() +
  facet_null() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_continuous() +
  theme_bw()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=gear), 
             stat = "identity") +
  coord_cartesian() +
  facet_null() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_continuous() +
  theme_classic()
```

```{r}
ggplot(data = mtcars) + 
  geom_point(mapping = aes(x = disp, y = hp, color=gear), 
             stat = "identity") +
  coord_cartesian() +
  facet_null() +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_continuous() +
  theme_minimal()
```