---
weight: 2
title: "Scatterplot"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## Import Data
```{r}
#---Loading data
load(url('https://github.com/MingChen0919/gst-colloquium-workshop-2018/raw/master/content/datasets/datasets.RData'))
```

## Load ggplot2 library
```{r}
#---loading ggplot library
library(ggplot2)
```

## Visualization

### Plot One; collective variables and macrostates
* Geom object: point
    * TIC-0 -> x
    * TIC-1 -> y
    * Macrostate -> color
  
```{r}
#---plot one
ggplot(energy_tics) +
  geom_point(mapping = aes(x = TIC0, y = TIC1, color = as.factor(macrostate)),
             stat = "identity", 
             position = "identity") +
  coord_cartesian() +
  facet_null() +
  
  ##----map data values to visual values
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_identity() +
  theme()
```

### Plot Two; manipulating axes, point colors
* Geom object: point
    * TIC-0 -> x
    * TIC-1 -> y
    * Macrostate -> color

* Plot edits
    * Set the axis labels (breaks)  
    * Set the axis limits (limits)  
    * Change axis orientation (trans = "reverse")  
    * Change colors for macrostates (scale_color_manual())  
    * Set labels for macrostates (breaks; labels)  
    * Change theme (theme_classic())  
```{r}
#---plot two
ggplot(energy_tics) +
  geom_point(mapping = aes(x = TIC0, y = TIC1, color = as.factor(macrostate)),
             stat = "identity", 
             position = "identity") +
  coord_cartesian() +
  facet_null() +
  
  ##----map data values to visual values
  scale_x_continuous(name = "TIC-0", 
                     breaks = c(0, 1, 2, 3, 4, 5, 6),
                     limits = c(0, 6)) +
  scale_y_continuous(name = "TIC-1",
                     breaks = -7:3, 
                     trans = "reverse",
                     limits = c(3, -7)) +
  scale_color_manual(values = c("1" = "red", "2" = "blue", "3" = "yellow", "4" = "green"),
                     name = "Macrostate", 
                     breaks = c(1,2,3,4),
                     labels = c('state 1', 'state 2', 'state 3', 'state 4')) + 
  theme_classic()
```

### Plot Three; add additional factor to visualize
* Geom object: point
    * TIC-0 -> x
    * TIC-1 -> y
    * Macrostate -> shape
    * Energy -> color
* Plot edits
    * Energy added as parameter
    * Change macrostate label
    
```{r}
#---plot three
ggplot(energy_tics) +
  geom_point(mapping = aes(x = TIC0, y = TIC1, shape = as.factor(macrostate), color = Energy),
             stat = "identity", position = "identity") +
  coord_cartesian() +
  facet_null() +
  
  ##----map data values to visual values
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_gradient() +
  scale_shape_discrete() +
  theme() +
  
  ##---change macrostate legend label
  labs(shape="Macrostate") 
```

### Plot Four; manipulate color gradients, specify point shapes
* Geom object: point
    * TIC-0 -> x
    * TIC-1 -> y
    * Macrostate -> shape
    * Energy -> color

* Plot edits
    * Change color gradient for energy (scale_color_gradient)
    * Specify shapes for macrostates
```{r}
#---plot four
ggplot(energy_tics) +
  geom_point(mapping = aes(x = TIC0, y = TIC1, shape = as.factor(macrostate), color = Energy),
             stat = "identity", position = "identity") +
  coord_cartesian() +
  facet_null() +
  
  ##----map data values to visual values
  scale_x_continuous(name = "TIC-0") +
  scale_y_continuous(name = "TIC-1",
                     breaks = -7:3) +
  scale_color_gradient(low="blue", high="red",
                       breaks = 0:8) +
  scale_shape_manual(values = c("1" = 1, "2" = 2, "3" = 3, "4" = 4),
                     name = "Macrostate") +
  theme_classic()
```

### Plot Five; add confidence ellipses
* Geom object: point
    * TIC-0 -> x
    * TIC-1 -> y
    * Macrostate -> shape
    * Energy -> color
* Geom object: path
    * TIC-0 -> x
    * TIC-1 -> y
    * Macrostate -> color
    * confidence ellipse -> stat
  
* Plot edits
    * Change legend labels (labs())
```{r}
#---plot five
ggplot(energy_tics) +
  geom_point(mapping = aes(x = TIC0, y = TIC1, fill = Energy, shape = as.factor(macrostate)),
             stat = "identity", position = "identity", inherit.aes = FALSE) +
  geom_path(mapping = aes(x = TIC0, y = TIC1, color = as.factor(macrostate)), 
            stat = "ellipse", inherit.aes = FALSE) + 
  coord_cartesian() +
  facet_null() +
  
  ##----map data values to visual values
  scale_x_continuous(name = "TIC-0") +
  scale_y_continuous(name = "TIC-1") +
  scale_shape_manual(values = c("1" = 21, "2" = 22, "3" = 23, "4" = 24)) +
  scale_color_discrete() +
  scale_fill_gradient(low = "blue", high = "red") +
  theme() +
  
  ##---change macrostate legend label
  labs(shape="Macrostate", color = "Macrostate") 
```
