---
weight: 2
title: "1/29/2018 Group Meeting Notes"
author: "Rachel Steele"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache.rebuild = TRUE)
```


## John presented his data and how it is set up within his .dat trajectory files:
* 29 Columns – representing features of molecular dynamics data
* 5000 rows – representing distances of carbons in protein backbone from one another at different time points

## Stuff we need to work on:
* A data parser – main problem seemed to be figuring out how to parse header column separately from data values, as delimiter is different for data values (space is delimiter) and column names (-- is delimiter)

Below is my solution for reference. (Ming): 
```{r}
library(ggplot2)
library(stringr)

# data path
data_path = '../datasets/John_datasets/traj_0.dat'

# get first line of the tray_0.dat and split by '--' to get all feature names
feature_names = str_split(readLines(data_path)[1], '--')[[1]]

# read data without header
data_raw = read.table(data_path, header = FALSE, sep = ' ', comment.char = 'D')

# assign feature names to the obtain data
colnames(data_raw) = feature_names

# show structure of the data, which is a data frame with 5001 rows and 29 columns.
str(data_raw)
```

* John -- make jupyter notebooks accessible to everyone and everyone else look at the jupyter notebooks -- this will give us a good baseline for figuring out what to do with the data in R (though John has it coded in python)
* To start creating different plots:
    - John had the idea of plotting frame values vs feature values (1-29)
        * To do this we will need to pull the y value from the header for labeling purposes as the numbers 1-29 are not very informative
        * solution for reference (MC):
            + see [GEOM_POINT for scatter plot](http://ggplot2.tidyverse.org/reference/geom_point.html)
            + labels are in the `feature_names` variable
            
* To figure out in R how to bin the data – so we can make a histogram of p-values indicating probability of certain values to occur
    + solution for reference (MC)
        * bin single value with histogram, see [GEOM_HISTOGRAM](http://ggplot2.tidyverse.org/reference/geom_histogram.html)
        * bin two variables with [GEOM_BIN2D](http://ggplot2.tidyverse.org/reference/geom_bin2d.html) or [GEOM_HEX](http://ggplot2.tidyverse.org/reference/geom_hex.html)
    
* Figure out how to make a covariance matrix using sliding model for x number of time steps as presented
    + covariance matrix example (MC)
  
```{r}
my_matrix = data.frame(x = sample(1:100, 10),
                       y = rnorm(10, 5, 10),
                       z = 1:10 + rnorm(1, 2))
my_matrix
cov(my_matrix)
```
  



<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>