---
weight: 1
title: "RStudio Accounts"
author: "gst data visualization group"
output: html_document
---

<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Workshop Environment

**RStudio Server Account**

Find your name in this [google sheet](https://docs.google.com/spreadsheets/d/1eR3eR62F1jozQHrZhTrnFEtRfVa1D6n52nXzPVIwfTA/edit#gid=0) and use the associated username and password to login to this [RStudio Server](http://149.165.157.82/).

We have a [second server](http://129.114.16.58/) running as a backup. Please use this one only when the first one does not work for you.

**Running a docker container locally**

Thanks to the machine learning workshop from last week! Everyone should have had docker installed. If you are interestd in running a local RStudio server. You can use the command below:

```{bash}
docker pull mingchen0919/gst-dv-rstudio
docker run --name=gst_dv -d -p 80:8787 mingchen0919/gst-dv-rstudio
```

You RStudio server should be available at http://127.0.0.1.




### Install packages

In case that you want to add your own packages, you can use the code below to install R packages in the RStudio server.

```{r}
dir.create('~/my-r-lib')
.libPaths('~/my-r-lib')
install.packages('ggplot2')
```


## Build Workshop Environment

If you are interested in how we set up the RStudio on the jetstream cloud computing environment, below is the instruction. 

### RStudio Docker on Jetstream

Launch a Jetstream image with Docker installed on Jetstream. Then run the following to luanch an RStudio server.

```{bash}
sudo su

# docker pull rocker/rstudio

# docker run --name=gst_dv -d -p 80:8787 rocker/rstudio

docker pull mingchen0919/gst-dv-rstudio
docker run --name=gst_dv -d -p 80:8787 mingchen0919/gst-dv-rstudio
```

### Create 30 new users

```
# login to the rstudio container
docker exec -it gst_dv bash

# create users in the rstudio container
for id in $(seq 1 45)
do
  echo "gst_user_${id}:passwd_user_${id}::::/home/gst_user_${id}:" | newusers
done
```

## For admin users

* [code 1](https://mingchen0919.github.io/gst-colloquium-workshop-2018/solutions/exercise_1)
* [code 2](https://mingchen0919.github.io/gst-colloquium-workshop-2018/solutions/tutorial3_codes)


