---
weight: 3
title: "ggplot2"
author: "gst data visualization group"
output: html_document
---

<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
options("width" = 200)
```

## ggplot2 graph template

```{r eval=FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>),
                  stat = <STAT>, position = <POSITION>) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION> +
  <SCALE_FUNCTION> +
  <THEME_FUNCTION>
```

* **GEOM_FUNCTION**: use a geometric object to represent your data.
    + ***MAPPINGS***: associate data values with visual values. e.g., categorial data -> different colors.
    + ***STAT***: use raw data values or calculate new values from raw values.
    + ***POSITION***: adjust overlapped geometric objects.
* **COORDINATE_FUNCTION**: type of coordinate systems.
* **FACET_FUNCTION**: divide plots into subplots
* **SCALE_FUNCTION**: how data values are mapped to visual values.
* **THEME_FUNCTION**: the apperance of graph components that is not directly related to the mappings.
