---
weight: 2
title: "Datasets"
output: html_document
---

<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
options("width" = 200)
```


```{r echo=FALSE}
load(url('https://github.com/MingChen0919/gst-colloquium-workshop-2018/raw/master/content/datasets/GST-WS.RData'))
```


## MD Features Trajectory

* Features Trajectory in `trajectory`
* The data format has 1 row per frame with N entries, and n rows

```{r}
trajectory[1:5, 1:3]
```

## Energy & TIC trajectory

* Use TICs as **collective variables**
    + Projection of a frame (741 features) onto each of d (a few) TIC (741x1 vector)
    + Reduces dimensionality 741→**d** of data
    + Capture informative (kinetic) data variation

```{r}
head(energy_tics[, 1:4])
```

## Microstate & Macrostate Trajectory:

* Microstate
  + After k-means clustering with k=1000
    - Each frame belongs to a cluster, ie microstate
    - Used TIC coordinate space (12 CVs)

* Macrostate
  + A Markov State Model was made first
    - Coarse-graining of microstates
    - Each frame now belongs to a macrostate
    - Used PCCA + manual

```{r}
head(energy_tics[, 5:6])
```

## Markov State Transition Matrix

* MSM: transition probability between (micro)states

    + Depends only on current state
    + Made by sliding window approach
    + Count microstate transitions from each i to each j, this ‘count’ matrix is the MSM

* Normalize counts: Probability of moving from state i to state j:  $t_{i,j}$
  
    + $t_{i,j=i}$ :  probability of staying in i
    + $t_{i,j \neq i}$ : probability of jumping to j
  
```{r}
protein_transition[1:5, 1:5]
```

## Stationary Distribution

* Equilibrium Population of States
    + First eigenvector of MSM transition matrix
    + Next eigenvectors are the kinetic processes
    + First eigenvector is 1
    + Next eigenvectors correspond to rates
  
```{r}
head(stationary_distribution)
```

