---
weight: 1
title: "Line plot for time series data"
output: html_document
---

<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Import data

```{r}
load(url('https://github.com/MingChen0919/gst-colloquium-workshop-2018/raw/master/content/S2-datasets/datasets.RData'))
```

## Load libraries

```{r}
library(ggplot2)
library(gridExtra)
```


## Visualization

### Feature against frame index

* **Geom object: line**
    + frame_index -> x
    + feature -> y

```{r}
# ggplot2 template
p1 = ggplot(data = trajectory) +
  geom_line(mapping = aes(x = frame_index/1000, y = trajectory[, 2]), 
            stat = "identity", position = "identity", color = '#ff6d6d') +
  coord_cartesian() +
  facet_null() +
  
  #----map data values to visual values---
  scale_x_continuous(name = "Frame index") +
  scale_y_continuous(name = colnames(trajectory[2])) +
  theme_classic()

p1
```

### Distribution of feature values

* **Geom object: density**
    + feature -> x
    
```{r}
# ggplot2 template
p2 = ggplot(data = trajectory) +
  geom_density(mapping = aes(x = trajectory[, 2]), 
            stat = "density", position = "identity", color = '#3dbfff') +
  coord_flip() +
  facet_null() +
  
  #----map data values to visual values---
  scale_x_continuous(name = colnames(trajectory)[2]) +
  theme_classic()
p2
```


### Combine time series plot and density plot

```{r}
grid.arrange(p1, p2, ncol = 2, widths = c(3,1))

# overwrite xlab on density plot.
grid.arrange(p1, p2 + xlab(''), ncol = 2, widths = c(3,1))
```



## All features in one plots

**Define your own plotting functions**

```{r}
line_plot = function(col_n) {
  p = ggplot(data = trajectory) +
    geom_line(mapping = aes(x = frame_index/1000, y = trajectory[, col_n]), 
              stat = "identity", position = "identity", color = '#ff6d6d') +
    coord_cartesian() +
    facet_null() +
    scale_x_continuous() +
    scale_y_continuous() +
    theme_classic() +
    theme(axis.title=element_text(size=5,face="bold")) +
    
    # helper functions
    xlab('Frame Index') +
    ylab(colnames(trajectory)[col_n])

  return(p)
}

density_plot = function(col_n) {
  p = ggplot(data = trajectory) +
    geom_density(mapping = aes(x = trajectory[, col_n]), 
              stat = "density", position = "identity", color = '#3dbfff') +
    coord_flip() +
    facet_null() +
    scale_x_continuous() +
    theme_classic() +
    theme(axis.title=element_text(size=5,face="bold")) +
    
    # helper functions
    xlab('')
    
  return(p)
}
```

Examples

```{r}
line_plot(2)
density_plot(4)
```

### Plots vector

```{r fig.height=20, fig.width=10, eval=TRUE}
all_plots = list()
for (i in 1:58) {
  
  # get column number
  col_n = (i + 1) %/% 2 + 1
  cat('The column number is: ', col_n, '\n')
  
  if (i %% 2 == 1) {
    cat('i is an odd number: i = ', i, '\n')
    all_plots[[i]] = line_plot(col_n)
  }
  
  else {
    cat('i is an even number: i = ', i, '\n')
    all_plots[[i]] = density_plot(col_n)
  }
  
  cat('----------------------------------\n')
}


all_plots$ncol = 6
all_plots$widths = c(5,2,5,2,5,2)
do.call(grid.arrange,  all_plots)
```

