---
weight: 3
title: "Network for graph-like data"
author: "GST Data Visualization Group"
output: html_document
---

<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Import data

```{r}
transition = read.table(url('https://raw.githubusercontent.com/MingChen0919/gst-colloquium-workshop-2018/master/content/datasets-2/John_datasets/transitionmatrix.dat'))
```


## Load libraries

```{r}
library(ggnetwork)
library(sna)
```


## Simple network

```{r}
n = network(transition)

n1 = ggnetwork(n)

n1$x = n1$x[, 1]
n1$y = n1$y[, 1]
n1$xend = n1$xend[, 1]
n1$yend = n1$yend[, 1]

ggplot(n1, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_nodes() +
  geom_edges() +
  theme_blank()
```

## Add attributes to **vertices**

Assign an importance level to each vertex

```{r}
n %v% "importance" = sample(1:6, nrow(transition), replace = TRUE)
```

Assign group to vertex

```{r}
n %v% "group" = sample(c("A", "B", "C"), nrow(transition), replace = TRUE)
```

Visualize **group** and **importance** informaction.

```{r}
n2 = ggnetwork(n)

n2$x = n2$x[, 1]
n2$y = n2$y[, 1]
n2$xend = n2$xend[, 1]
n2$yend = n2$yend[, 1]


ggplot(n2, mapping = aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_nodes(mapping = aes(color = group, size = importance)) +
  geom_edges() +
  theme_blank()
```

## Adjust edges

Edge transprancy

```{r}
ggplot(n2, mapping = aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_nodes(mapping = aes(color = group, size = importance)) +
  geom_edges(alpha = 0.05) +
  theme_blank()
```

Threshold for edge display

```{r}
threshold = 0.1
transition2 = transition
transition2[transition2 < 0.1] = 0


n = network(transition2)

n %v% "importance" = sample(1:6, nrow(transition), replace = TRUE)
n %v% "group" = sample(c("A", "B", "C"), nrow(transition), replace = TRUE)

n1 = ggnetwork(n)

n1$x = n1$x[, 1]
n1$y = n1$y[, 1]
n1$xend = n1$xend[, 1]
n1$yend = n1$yend[, 1]

ggplot(n1, mapping = aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_nodes(mapping = aes(color = group, size = importance)) +
  geom_edges(alpha = 0.8) +
  theme_blank()
```

## Add labels to nodes

```{r}
ggplot(n1, mapping = aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(alpha = 0.8) +
  geom_nodelabel(aes(label = vertex.names, color = as.factor(importance)), 
                 cex = 2, 
                 fontface = "bold") +
  scale_color_manual(name = "Importance", values = colorRampPalette(c("yellow", "red"))(6)) +
  theme_blank()
```

